---
title: "What even is Hyperpop?"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    storyboard: true
---

```{r setup, include=FALSE}
library(flexdashboard)
```

-----------------------------------------------------------------------

### Introduction

```{r}

```
##### Corpus: 
The following spotify editorial playslits [hyperpop](https://open.spotify.com/playlist/37i9dQZF1DX7HOk71GPfSw), [hyperpop classics](https://open.spotify.com/playlist/37i9dQZF1DXakZCf7SV1CS) and [feedBack](https://open.spotify.com/playlist/37i9dQZF1DXbmErmrio5En). Other similar lists: [misfits 2.0](https://open.spotify.com/playlist/37i9dQZF1DXciq74vrQmM8) and [Lorem](https://open.spotify.com/playlist/37i9dQZF1DXdwmD5Q7Gxah) (will not be analyzed)

##### Some question ideas: (WIP)
Are they all just the same? How do they compare to other "genre-based" playlists curated by the artists themselves instead? (e.g. phritz's [Botanica](https://open.spotify.com/playlist/5KgQT9YWz3EqhIN4Jh69IU?si=dd13155eb2dc4c80), underscores's [arson tutorial](https://open.spotify.com/playlist/4yx17NWwtN8Ij7zHQFrX9c?si=a7082850d1614cdd), jane remover's [Required Listening---](https://open.spotify.com/playlist/3B2fWxsJHxygGh7toOS09u?si=b1bc739e0ba24a3f) etc.)

##### Description of the corpus
These playlists usually contain songs curated by the editorial team in Spotify that dictate the listening habits of thousands of users and very importantly, the artist's (very frequently small, independent, queer and POC) revenue and streaming stats (e.g. monthly listeners, playlists placements, streams per song, the "fans also like" section, etc). Seeing that they carry such an important role , I think it is important to study them and try to understand them. To the average indie artist, these playlists are a to-be-solved puzzle that has the power to dictate their music career. It's a never ending fight against trends, algoritgms, and of course, creative drive and expression. From studying them, I expect to find common features that connect songs directly (features, genre, instruments, BPM) and more indirectly (tone of lyrics/instrumental, aesthetic/vibe, artist's network (like their friends, colleges, production/writing creadits, etc). I expect these common features to be decesive in determining if a song gets chosen by a spotify employee to be part of one of these playlists. I am not so sure if algorithms recommendations will play a very important roles in editorial playlist, (I know they do in algorithmic-based playlists) but i think it can also be another intersting topic to study.

The word "hyperpop" has a life of its own at this point. Made up by a conoction of songs created by Spotify in 2019, the hyperpop has been said to encapsulate a very specific sound<sub>1</sub>. But does it really? It is clearly not a genre, because the contents of the playlist are so incredibly broad and creatively diverse, that labeling it that way would feel like an understatement<sub>2</sub> (in my opinion). The sound has shifted, trends have changed, and the label can't be used as a genre for a sound. Well, I guess you can use it if you want, but I think It would be rather unprecise. Can we see any specific musical characteristics throughout that repeat across the songs in these online lists? What of these common charactersitics are "enough" to call something a genre?

What many people would originally call "[hyperpop](https://open.spotify.com/playlist/37i9dQZF1DX7HOk71GPfSw)" in 2020 is now under the playlist "[hyperpop classics](https://open.spotify.com/playlist/37i9dQZF1DXakZCf7SV1CS)", which mainly contains artists from the realms of PC-Music and co. (such as SOPHIE - Immaterial, 100 gecs - Hand crushed by a mallet, A. G. Cook - superstar) but also artists from smaller scenes from soundclound (like underscores - second hand embarassment, quinn - from paris, with love, d0llywood1 - ihonestlymightjustgiveup, brakence - rosier/punk2, etc). Nowadays, the hyperpop playlist is home to select artist from that cohort, but also welcomes new faces (such as vai5000 - replicant, tsubi club - burbank house, Musa - red). It is clear that the "classic" essence is still there, something has changed, and the "hyperpop sound" is now getting mixed with new ideas. In some cases, the genre has even been bent to an extent that leaves the classic "hyperpop" sound. These new artists are now also included in the new "[feedBack](https://open.spotify.com/playlist/37i9dQZF1DXbmErmrio5En)" playlist, homing previously known faces such as underscores - Count of three and brakence - 5g plus some new ideas that are clearly inspired by the sound of the incredibly broad umbrella term (defsharp - offensive mechanism and gabby start - welcome to the plant cell bath). With such an extensive and ever-growing scene, is it fair to let big corporations like Spotify decide what makes a genre or what doesn't?<sub>2</sub>

##### Footnotes
1. The "classic hyperpop sound" is characterized by complex, hard-hitting and jaw-dropping sound design, catchy high-pitched vocals and blaring melodies with distorted bass patterns.
2. See ["why hyperpop and digicore don't exist according to SoundCloud's creative director"](https://www.youtube.com/watch?v=HGzP1naK_sU), an interview with Billie Bugara, creator of the indie label [DeadAir](https://www.billiebugara.com/deadair#:~:text=deadAir%20Records%20was%20co%2Dfounded,their%20artistic%20careers%20take%20shape.)


-----------------------------------------------------------------------

### The "hyperpop sound"? 

<p>I created this plot because I wanted to compare the main attributes that apply to most "hyperpop music". The plot is divided in four, one for each playlist. In this case, I studied the original "hyperpop" playlist, the "hyperpop classics" playlist (both editorial spotify playlist) but I also included the "sleepy.picks" playlist and the "Perfect Music Friday" playlist. The latter two are also curated, but in this case by artists in the scene. I thought it would be interesting to see if the tracks that Spotify employees select have similar attributes to what the actual artists in the scene pick.</p>


```{r}
library(tidyverse)
library(spotifyr)
library(plotly)

hyper <- get_playlist_audio_features("", "37i9dQZF1DX7HOk71GPfSw")
classic <- get_playlist_audio_features("", "37i9dQZF1DXakZCf7SV1CS")
sleepy <- get_playlist_audio_features("", "1darBgHtPFeoXVS1qvXhUD")
perfect <- get_playlist_audio_features("", "6oaHLB9qwLOo0RVpU2gGJ5")

corpus <-
  bind_rows(
    hyper |> mutate(category = "Hyperpop"),
    classic |> mutate(category = "Hyperpop Classic"),
    sleepy |> mutate(category = "Sleepy Zone by Blisscord"),
    perfect |> mutate(category = "Perfect Music Friday by PC Music"),
  )

corpus |>
  ggplot(aes(x = valence, y = energy, size = tempo, color = category)) +
  geom_jitter(alpha = 0.8,) +
  facet_wrap(~category) +
  scale_color_brewer(
    type = "qual",
    palette = "Spectral",
    guide = "none",
  ) +
  scale_size_continuous(
    
    labels = waiver(),
    name ="BPM"

  )

ggplotly(corpus)
```
<p>I chose to display energy because it measures if tracks feel fast, loud, and noisy. I contemplated using loudness instead but I thought energy would provide more information and it would be easier to understand than a dB scale with loudness. I also measured the valence (how happy or sad the songs are) and BPM. </p>

<p>With this plot, I found that the bulk of tracks in the editorial playlists were very energetic and happy. Moreover, "hyperpop" playlist only included 6 songs under 0.25 valence, one of which had very low BPM. In comparison, the songs included in the playlists made by curators inside the community (mainly artists) like the sleepy playlist and the Perfect Music Friday playlists are more spread out in terms of energy and valence. One thing that seems to be consistent though is the overall BPM, that tends to stay very high across all playlists.</p>

### Discussion

```{r}

```

